---
title: "GSS Challenge More Data"
author: "Tyler Hamblin"
date: "28 September 2023"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom?


$H_A$: A change in race statistically results in a change in opinion towards capitol punishment. 


$H_0$: The two factors race and cappun have no affect on each other. 



```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The CandP helper files are going to assist us in comparing the two data variables of cappun and race. Due to this being a Categorical ~ Categorical question we will first use a bar chart, then we will run a favstats command to give us graphical and numerical results. 

Under the inferential results section we will run a Chi-Squared test which will suggest if there is a statistical difference in the data or not. To figure out where the idfference is, if there is one, we could normally run a fisher test but in this case there are more than two factors under the "race" category so the fisherXY function will replace it to be sufficient for all combinations. 


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```


According to the data there is a clear skew of white people that are in favor of capitol punishment when compared to other races. This is not continuous through out the other races and it is typical for black people especially to oppose capitol punishment. 


## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

As seen in the numerical breakdown of the data, about twice the percentage of white people are in favor of capitol punishment. When looking at black and other races there is about a 50:50 spread of whether or not an individual from that category will be in support or not. 



## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
```


By running a Chi-Squared test we can prove that the Chi-Squared Statistic is higher than the critical value derived from our degrees of freedom. This big number shows that there is significant difference in the data when compared to an expected value. A very small p-value, $2.2 \text{ x } 10^{-16}$, suggests that there is a near guarantee that there will be a difference somewhere along the data. The exact location can be found with a Fisher test. 



```{r}
fisherXY(mydata$race,mydata$cappun,printAll = TRUE)
```


By running a FisherXY test with a confidence level of 99% we can see where the p-value suggests a difference in data occurs. the relationship between the black and white races is where the greatest significance can be noticed. It is also notable that almost all of the p-values are less than 0.05 but one is significantly lower than the rest and deserves extra attention. 


## More data version


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

By using graphical and numerical strings of code, the white race typically has more instances of being in favor of capitol punishment when compared to the other race categories. After determining that a statistical difference occurred somewhere and solidifying the evidence between the Black and White races, we can reject the null hypothesis and accept the alternative. In addition we used branch 2 to look into a similar circumstance but with a smaller sample size. We have similar results after running the bar graph and percentage numerical results commands. The multiple p-values in this branch are replaced by a single p-value in the second and a smaller Chi-Squared Statistic, from 80 to 19, is used. What can be interpreted from this is that the larger the data frame, the more differences appear. We can still state that there is correlation in the two cases but more 
research would be needed to decided what the cause of this change is. 




