---
title: "GSS Challenge Less Data"
author: "Tyler Hamblin"
date: "29 September 2023"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
source(file="fisherXY.R")

GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??



$H_A$: A change in race statistically results in a change in opinion towards capitol punishment. 


$H_0$: The two factors race and cappun have no affect on each other. 



```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


By looking at the data in this project we can see that the data is Categorical~Categorical meaning that we will use the Cat_Cat helper file from the CandP zip. We will run commands for a bar chart to graphically express the data and a table to show the numerical version. Finally under the inferential data section a Chi-Squared test will tell us if there is a statistical difference or not but not where the difference is. For that we will rely on a FisherXY test which runs the fisher test in a loop for a multitude of combinations. 


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results




### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->




```{r}
ggplot(mydata, aes(x = race, fill = cappun)) +
  geom_bar(position = "dodge") +
  labs(title = "Capital Punishment By Race",
       x = "Race",
       y = "Count") +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5,size = 18))
```

Again we see the same disproportionate ammount of white peopel in favor of capitol punishment. Black people in this data set still are slightly more opposed to the use of it. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

Another near 50:50 split can be seen with the black and other race categories. White people have about twice as many people in support when compared to those who oppose. 




## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

$\Huge\chi^2=\sum\frac{(O-E)^2}{O}$


From the Chi-Squared test we see that the Statistic is still much higher than the critical value provided by the degrees of freedom. The large number shows there is a difference from the expected value and a small p-value shows statistical variation as well. A fisher test needs to be ran in order to find out where the difference occurs amongst the data. 


```{r ,eval=FALSE, include=TRUE}
FisherTable(table1,confL = 0.95,printAll = TRUE)
```


```{r}
sub_whiteBlack <- mydata %>% filter(race %in% c("white", "black"))
sub_whiteBlack$race <- factor(sub_whiteBlack$race, levels = c("white", "black") )
table2 <- xtabs(~cappun + race, data=sub_whiteBlack)
fisherResult <- fisher.test(table2)
fisherResult
low=(fisherResult$conf.int[1]/(fisherResult$conf.int[1] + 1))*100
hi=(fisherResult$conf.int[2]/(fisherResult$conf.int[2] + 1))*100
print("Confidence Interval")
low
hi
```



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
